



# 1. `make_fd()` — the simulation function

Takes a matrix of coefficient draws and builds three counterfactual datasets: one where
every voter sits exactly at the reference party's position on the chosen dimension, and two
where they sit `delta` units either side of it (clamped to the 0–10 scale when
`squish = TRUE`).

For each draw it recomputes fitted probabilities without re-estimating the
model, differences them against the baseline, averages the two sides, and returns the mean
change in each party's vote probability.

```{r make-fd}
make_fd <- function(B,
                    model,
                    data,
                    ref_party = c("Reform", "Labour", "LibDem", "Conservative", "Green"),
                    dimension = c("tax", "ineq"),
                    delta = 1,
                    squish = TRUE,
                    ...){
  rp <- match.arg(ref_party)
  dim <- match.arg(dimension)

  if(!is.matrix(B))B <- matrix(B, nrow=1)
  tmp <- data
  ref <- subset(tmp, alt == rp) |> unclass()
  ref$chid <- unfold_idx(ref$idx)$chid
  ref <- ref[c("rp_tax", "rp_ineq", "chid")]
  ref <- do.call(data.frame, ref)
  names(ref) <- c("ref_tax", "ref_ineq", "chid")
  tmp <- left_join(tmp, ref, by = join_by(chid))

  tmp0 <- tmpm <- tmpp <- tmp
  tmp0[[paste0("lrself_", dim)]] <- tmp0[[paste0("ref_", dim)]]
  if(squish){
    tmpm[[paste0("lrself_", dim)]] <- ifelse(
      tmp0[[paste0("ref_",dim)]] - delta < 0,
      0,
      tmp0[[paste0("ref_",dim)]] - delta)
    tmpp[[paste0("lrself_", dim)]] <- ifelse(
      tmpp[[paste0("ref_",dim)]] + delta > 10,
      10,
      tmpp[[paste0("ref_",dim)]] + delta)
  }else{
    tmpm[[paste0("lrself_", dim)]] <- tmp0[[paste0("ref_",dim)]] - delta
    tmpp[[paste0("lrself_", dim)]] <- tmpp[[paste0("ref_",dim)]] + delta

  }

  res <- lapply(1:nrow(B), \(i){
    p0 <- update(model, data = tmp0, start = B[i,], iterlim = 0, print.level = 0)$probabilities
    pm <- update(model, data = tmpp, start = B[i,], iterlim = 0, print.level = 0)$probabilities
    pp <- update(model, data = tmpm, start = B[i,], iterlim = 0, print.level = 0)$probabilities
    diffm <- pm-p0
    diffp <- pp-p0
    diff <- (diffm + diffp)/2
    do.call(data.frame, as.list(colMeans(diff)))
  })
  do.call(rbind, res)
}
```


```{r packages-data}
library(tidyverse)
library(mlogit)
library(MASS)
library(dplyr)
library(ggplot2)
library(forcats)
library(here)
library(readr)
library(marginaleffects)
library(dfidx)

select <- dplyr::select

dat <- read_csv(here("data", "bes2024_small_clean.csv"))

dat <- dat %>%
  mutate(
    vote     = factor(vote, levels = c("Labour", "Conservative", "LibDem", "Reform", "Green")),
    age_band = factor(age_band),
    degree   = factor(degree),
    Region_lab = factor(Region_lab)
  )
```

# 3. Reshape to long format

Identifies the `rp_tax.*` and `rp_ineq.*` columns holding each party's position, then expands each
respondent into five rows, one per party alternative, with the party positions as
alternative-specific variables.

```{r reshape}
tax_cols  <- grep("^rp_tax\\.",  names(dat))
ineq_cols <- grep("^rp_ineq\\.", names(dat))

dat2 <- dat %>%
  dplyr::select(vote, Age, lrself_tax, lrself_ineq,
                age_band, degree, Region_lab, wt_vote,
                all_of(names(dat)[tax_cols]),
                all_of(names(dat)[ineq_cols])) %>%
  na.omit() %>%
  as.data.frame()

mldat <- mlogit.data(dat2, choice = "vote", shape = "wide",
                     varying = c(which(names(dat2) %in% names(dat)[tax_cols]),
                                 which(names(dat2) %in% names(dat)[ineq_cols])))
```

# 4. The models

`mod_party` is a conditional logit where Age and degree get party-specific coefficients and
the two absolute distances (tax, inequality) get a **separate proximity coefficient for each
party**, so proximity can matter more for some parties than others. `mod_party2` fits the
same variables with a single shared proximity coefficient.

```{r models}
mod_party <- mlogit(
  vote ~ 0 | Age + degree | I(abs(lrself_tax - rp_tax)) + I(abs(lrself_ineq - rp_ineq)) ,
  data = mldat, reflevel = "Labour",
  weights = wt_vote
)

mod_party2 <- mlogit(
  vote ~ I(abs(lrself_tax - rp_tax)) + I(abs(lrself_ineq - rp_ineq)) | Age + degree ,
  data = mldat, reflevel = "Labour",
  weights = wt_vote
)
```



# 5. Coefficient draws

Draws 1,000 coefficient vectors from a multivariate normal centred on the estimates with the
model's covariance matrix — a parametric bootstrap used to put confidence intervals on the
simulated quantities. 

```{r draws}
library(future)
library(future.apply)
plan(multisession(workers = 10))

B <- MASS::mvrnorm(1000, coef(mod_party), vcov(mod_party))
Blist <- asplit(B, 1)
```

# 6. Running the full grid

Repeats the same block ten times: five reference parties crossed with the two dimensions,
each looped over delta values 1 to 5, running `make_fd` across all 1,000 draws. Each block
keeps only the column for its own reference party — i.e. what happens to Reform's vote
probability as voters move away from Reform — accumulating into `all_res_tax` and then
`all_res_ineq`.

## 6a. Tax dimension

```{r grid-tax-reform}
res <- lapply(1:5, \(i)future_lapply(
  Blist,
  make_fd,
  model = mod_party,
  data = mldat,
  ref_party = "Reform",
  delta=i,
  future.packages = c("mlogit", "dplyr"),
  future.seed = TRUE
))

res <- lapply(res, \(x)bind_rows(x))
res <- bind_rows(res, .id = "delta") |>
  mutate(delta = as.numeric(delta))
all_res <- NULL
all_res <- all_res |>
  bind_rows(
    res |> select(delta, Reform) |> mutate(party = "Reform") |> rename(vote_prob = Reform)
  )
```

```{r grid-tax-conservative}
res <- lapply(1:5, \(i)future_lapply(
  Blist,
  make_fd,
  model = mod_party,
  data = mldat,
  ref_party = "Conservative",
  delta=i,
  future.packages = c("mlogit", "dplyr"),
  future.seed = TRUE
))

res <- lapply(res, \(x)bind_rows(x))
res <- bind_rows(res, .id = "delta") |>
  mutate(delta = as.numeric(delta))

all_res <- all_res |>
  bind_rows(
    res |> select(delta, Conservative) |> mutate(party = "Conservative") |> rename(vote_prob = Conservative)
  )
```

```{r grid-tax-labour}
res <- lapply(1:5, \(i)future_lapply(
  Blist,
  make_fd,
  model = mod_party,
  data = mldat,
  ref_party = "Labour",
  delta=i,
  future.packages = c("mlogit", "dplyr"),
  future.seed = TRUE
))

res <- lapply(res, \(x)bind_rows(x))
res <- bind_rows(res, .id = "delta") |>
  mutate(delta = as.numeric(delta))

all_res <- all_res |>
  bind_rows(
    res |> select(delta, Labour) |> mutate(party = "Labour") |> rename(vote_prob = Labour)
  )
```

```{r grid-tax-libdem}
res <- lapply(1:5, \(i)future_lapply(
  Blist,
  make_fd,
  model = mod_party,
  data = mldat,
  ref_party = "LibDem",
  delta=i,
  future.packages = c("mlogit", "dplyr"),
  future.seed = TRUE
))

res <- lapply(res, \(x)bind_rows(x))
res <- bind_rows(res, .id = "delta") |>
  mutate(delta = as.numeric(delta))

all_res <- all_res |>
  bind_rows(
    res |> select(delta, LibDem) |> mutate(party = "LibDem") |> rename(vote_prob = LibDem)
  )
```

```{r grid-tax-green}
res <- lapply(1:5, \(i)future_lapply(
  Blist,
  make_fd,
  model = mod_party,
  data = mldat,
  ref_party = "Green",
  delta=i,
  future.packages = c("mlogit", "dplyr"),
  future.seed = TRUE
))

res <- lapply(res, \(x)bind_rows(x))
res <- bind_rows(res, .id = "delta") |>
  mutate(delta = as.numeric(delta))

all_res <- all_res |>
  bind_rows(
    res |> select(delta, Green) |> mutate(party = "Green") |> rename(vote_prob = Green)
  )

all_res_tax <- all_res
```

## 6b. Inequality dimension

Identical to the above, with `dimension = "ineq"` passed through to `make_fd` so the
inequality/redistribution scale is the one being run `all_res` is reset to `NULL`
first and snapshotted as `all_res_ineq` at the end.

```{r grid-ineq-reform}
res <- lapply(1:5, \(i)future_lapply(
  Blist,
  make_fd,
  model = mod_party,
  data = mldat,
  dimension = "ineq",
  ref_party = "Reform",
  delta=i,
  future.packages = c("mlogit", "dplyr"),
  future.seed = TRUE
))

res <- lapply(res, \(x)bind_rows(x))
res <- bind_rows(res, .id = "delta") |>
  mutate(delta = as.numeric(delta))
all_res <- NULL
all_res <- all_res |>
  bind_rows(
    res |> select(delta, Reform) |> mutate(party = "Reform") |> rename(vote_prob = Reform)
  )
```

```{r grid-ineq-conservative}
res <- lapply(1:5, \(i)future_lapply(
  Blist,
  make_fd,
  model = mod_party,
  data = mldat,
  dimension = "ineq",
  ref_party = "Conservative",
  delta=i,
  future.packages = c("mlogit", "dplyr"),
  future.seed = TRUE
))

res <- lapply(res, \(x)bind_rows(x))
res <- bind_rows(res, .id = "delta") |>
  mutate(delta = as.numeric(delta))

all_res <- all_res |>
  bind_rows(
    res |> select(delta, Conservative) |> mutate(party = "Conservative") |> rename(vote_prob = Conservative)
  )
```

```{r grid-ineq-labour}
res <- lapply(1:5, \(i)future_lapply(
  Blist,
  make_fd,
  model = mod_party,
  data = mldat,
  dimension = "ineq",
  ref_party = "Labour",
  delta=i,
  future.packages = c("mlogit", "dplyr"),
  future.seed = TRUE
))

res <- lapply(res, \(x)bind_rows(x))
res <- bind_rows(res, .id = "delta") |>
  mutate(delta = as.numeric(delta))

all_res <- all_res |>
  bind_rows(
    res |> select(delta, Labour) |> mutate(party = "Labour") |> rename(vote_prob = Labour)
  )
```

```{r grid-ineq-libdem}
res <- lapply(1:5, \(i)future_lapply(
  Blist,
  make_fd,
  model = mod_party,
  data = mldat,
  dimension = "ineq",
  ref_party = "LibDem",
  delta=i,
  future.packages = c("mlogit", "dplyr"),
  future.seed = TRUE
))

res <- lapply(res, \(x)bind_rows(x))
res <- bind_rows(res, .id = "delta") |>
  mutate(delta = as.numeric(delta))

all_res <- all_res |>
  bind_rows(
    res |> select(delta, LibDem) |> mutate(party = "LibDem") |> rename(vote_prob = LibDem)
  )
```

```{r grid-ineq-green}
res <- lapply(1:5, \(i)future_lapply(
  Blist,
  make_fd,
  model = mod_party,
  data = mldat,
  dimension = "ineq",
  ref_party = "Green",
  delta=i,
  future.packages = c("mlogit", "dplyr"),
  future.seed = TRUE
))

res <- lapply(res, \(x)bind_rows(x))
res <- bind_rows(res, .id = "delta") |>
  mutate(delta = as.numeric(delta))

all_res <- all_res |>
  bind_rows(
    res |> select(delta, Green) |> mutate(party = "Green") |> rename(vote_prob = Green)
  )

all_res_ineq <- all_res
```


# 7. Draft plots

Collapses the draws into a posterior mean and 2.5/97.5 percentile interval for each delta and
party. Plots these as a line  faceted by party, separately for tax and
inequality.

```{r draft-plots}
all_res_tax |>
  summarise(est = mean(vote_prob),
         lwr = quantile(vote_prob, .025),
         upr = quantile(vote_prob, .975),
         .by = c(delta, party)) |>
  ggplot(aes(x=delta, y = est, ymin = lwr, ymax=upr)) +
  geom_ribbon(alpha=.25) +
  geom_line() +
  facet_wrap(~party) +
  theme_bw()

all_res_ineq |>
  summarise(est = mean(vote_prob),
            lwr = quantile(vote_prob, .025),
            upr = quantile(vote_prob, .975),
            .by = c(delta, party)) |>
  ggplot(aes(x=delta, y = est, ymin = lwr, ymax=upr)) +
  geom_ribbon(alpha=.25) +
  geom_line() +
  facet_wrap(~party) +
  theme_bw()
```

# 8. Party position map

Computes mean voter self-placement and mean party positions on both scales, then reshapes
them into one row per party. Plots them as a labelled scatter showing where parties and the
average voter sit in the two-dimensional economic space.

```{r position-map}
means <- dat2 |>
  summarise(across(c(starts_with("lrself"), starts_with("rp")), mean))

names(means) <- gsub("rp_(.*)\\.(.*)", "\\2_\\1", names(means))
names(means) <- gsub("lrself", "Voters", names(means))

means <- pivot_longer(means, cols=everything(), names_pattern = "(.*)_(.*)", names_to = c("party", ".value"))

library(ggrepel)
ggplot(means, aes(x=tax, y=ineq)) +
  geom_point() +
  geom_text_repel(aes(label = party)) +
  theme_minimal() +
  labs(x="Taxes", y="Inequality")
```

# 9. Pairwise significance tests

Reshapes the `delta == 1` draws into a draws-by-party matrix and multiplies it by a contrast
matrix covering all ten party pairs, giving a distribution of differences for each pair.
Converts the tail proportions into two-sided empirical p-values and filters to pairs
significant at p < .05.

```{r pairwise}
res_mat <- all_res |>
  filter(delta == 1) |>
  mutate(obs = row_number(), .by=party) |>
  pivot_wider(names_from = "party", values_from = "vote_prob") |>
  select(-c(delta, obs)) |>
  as.matrix()

combs <- combn(ncol(res_mat), 2)
D <- matrix(0, nrow=ncol(res_mat), ncol = ncol(combs))
D[cbind(combs[1,], 1:ncol(combs))] <- -1
D[cbind(combs[2,], 1:ncol(combs))] <- 1

pd <- res_mat %*% D
pdiff <- apply(pd, 2, \(x)mean(x < 0))
pdiff <- ifelse(pdiff > .5, 1-pdiff, pdiff)
tibble(
  smaller = colnames(res_mat)[combs[1,]],
  larger = colnames(res_mat)[combs[2,]],
  diff = colMeans(pd),
  pval = pdiff
) |>
  filter(pval < .05)
```



# 10. Final figures

Refactors the earlier plotting code into `summarise_delta_curve()` and `plot_delta_curve()`,
adding percent-formatted axes

```{r final-plots}
summarise_delta_curve <- function(df) {
  df |>
    summarise(est = mean(vote_prob),
              lwr = quantile(vote_prob, .025),
              upr = quantile(vote_prob, .975),
              .by = c(delta, party))
}

plot_delta_curve <- function(summary_df, dim_label) {
  ggplot(summary_df, aes(x = delta, y = est, ymin = lwr, ymax = upr)) +
    geom_ribbon(alpha = .25, colour = "transparent", fill = "#2a78d6") +
    geom_line(colour = "#2a78d6", linewidth = 1) +
    facet_wrap(~party, ncol = 3) +
    scale_y_continuous(labels = scales::percent) +
    labs(x = "Voter's distance from party",
         y = "Change in Pr(vote), vs. distance 0",
         title = dim_label) +
    theme_bw() +
    theme(panel.grid.minor = element_blank())
}

delta_curve_tax  <- summarise_delta_curve(all_res_tax)
delta_curve_ineq <- summarise_delta_curve(all_res_ineq)

p_delta_tax  <- plot_delta_curve(delta_curve_tax,  "Tax and spend")
p_delta_ineq <- plot_delta_curve(delta_curve_ineq, "Inequality / redistribution")

p_delta_tax
p_delta_ineq
```

# 11. Saving output



```{r save-tables}
# --- save tables (summarised, for plotting/reporting) -----------------

write.csv(delta_curve_tax,  here("tables", "delta_curve_tax.csv"),  row.names = FALSE)
write.csv(delta_curve_ineq, here("tables", "delta_curve_ineq.csv"), row.names = FALSE)

saveRDS(delta_curve_tax,  here("tables", "delta_curve_tax.rds"))
saveRDS(delta_curve_ineq, here("tables", "delta_curve_ineq.rds"))


# --- save underlying per-draw results too (raw, in case you need to
#     re-derive different summaries later without re-running the sims)

write.csv(all_res_tax,  here("tables", "all_res_tax.csv"),  row.names = FALSE)
write.csv(all_res_ineq, here("tables", "all_res_ineq.csv"), row.names = FALSE)

saveRDS(all_res_tax,  here("tables", "all_res_tax.rds"))
saveRDS(all_res_ineq, here("tables", "all_res_ineq.rds"))
```

```{r save-plots}
# --- save plots ---------------------------------------------------------

ggsave(here("plots", "delta_curve_tax.png"),  p_delta_tax,  width = 9, height = 6, dpi = 300)
ggsave(here("plots", "delta_curve_ineq.png"), p_delta_ineq, width = 9, height = 6, dpi = 300)
```
